### 2007-01-03 Documentation/Manual.tex
PH channged credits
Part of manual updated

\newcommand{\glbxxx}{hep-ph/0407333}
\newcommand{\capdef}{}
\newcommand{\mycaption}[2][\capdef]{\renewcommand{\capdef}{#2}%
        \caption[#1]{{ %\itshape
 {\footnotesize #2}}}}
\makeatletter
\renewcommand{\fnum@table}{\textbf{\tablename~\thetable}}
\renewcommand{\fnum@figure}{\textbf{\figurename~\thefigure}}
\makeatother

@@

\begin{center}
{\large Patrick Huber$^a$, Manfred Lindner$^b$, Walter Winter$^c$}
\end{center}

@@

\begin{center}
Version from \today\ for \GLOBES\ 2.0.11
\end{center}

@@

{\em\small $^{\mathrm{a}}$%
       Physics Department,
       University of Wisconsin,
       1150 University Av., Madison, WI 53706, USA}

{\em\small $^{\mathrm{b}}$%
       Institut f\"ur Theoretische Physik, Physik--Department,\\
       Technische Universit\"at M\"unchen,
       James--Franck--Strasse, D--85748 Garching, Germany}

{\em\small $^{\mathrm{c}}$%
       Institute for Advanced Study, 
       Einstein Drive, Princeton, NJ 08540, USA}

@@

\clearpage
\thispagestyle{empty}
\bigskip
\begin{quote}
    Copyright \copyright  2004,2005  The GLoBES Team.
    Permission is granted to copy, distribute and/or modify this document
    under the terms of the GNU Free Documentation License, Version 1.2
    or any later version published by the Free Software Foundation;
    with the invariant Sections ``Terms of usage of \GLOBES" 
    and ``Acknowledgments'', no Front-Cover Texts, and no Back-Cover Texts.
    A copy of the license is included in the section entitled "GNU
    Free Documentation License".
\end{quote}
\bigskip

@@

On the experiment definition side, either built-in neutrino fluxes
(\eg, neutrino factory) or arbitrary fluxes can be used. Similarly,
arbitrary cross sections, energy dependent efficiencies, the
energy resolution function, the considered oscillation channels, 
backgrounds, and many other features can be specified. 
For the systematics, energy
normalization and calibration errors can be simulated. Note that
the energy ranges and windows, as well as the bin widths can be
(almost) arbitrarily chosen, which means that variable bin
widths are allowed. Together with \GLOBES\ comes a number of
pre-defined experiments in order to demonstrate the capabilities
of \GLOBES\ and to provide prototypes for new experiments.

@@

\GLOBES\ is developed for academic use. Thus, the \GLOBES\ Team would
appreciate being given academic credit for it. Whenever you use \GLOBES\
to produce a publication or a talk indicate that you have used \GLOBES\ and
please cite the reference~\cite{globes_paper}
\begin{quote}
P. Huber, M. Lindner and W. Winter\\
Simulation of long baseline neutrino oscillation experiments with \GLOBES\\
arXiv:\glbxxx.  
\end{quote}
but \emph{not} this manual. This manual itself is not a scientific 
publication and will not be submitted to a scientific journal. 
It will evolve during time since it is intended for 
regular revision. Besides that, many of the data which are used by \GLOBES\ 
and distributed together with it should be properly referenced. 
For details see below.

@@

\index{norm}{Referencing!data in \GLOBES}
\GLOBES\ wouldn't be useful without having high quality input data.
Much of these input data have been published elsewhere and the authors
of those publications would appreciate to be cited whenever their work
is used. It is solely the user's responsibility 
to make sure that he understands where the input material for \GLOBES\ comes
from and if additional work has to be cited in addition to the 
\GLOBES\ paper~\cite{globes_paper}. To assist with this task, we provide  the necessary information for the data coming along together with \GLOBES.

@@

@@

The application 
software is, except from some example files, not part of \GLOBES , since
the evaluation of the experiment performance is often a matter of taste
and definition. In addition, the algorithms depend, especially for
high-precision instruments, very much on the oscillation parameters.
In general, it is quite simple to simulate superbeams and reactor
experiments. However, because of the more complicated topology, the
simulation of neutrino factories is much more difficult. In order
to demonstrate some of these difficulties, we present in this manual only
examples with neutrino factories. These examples can be found in
\Part~\ref{part:1} within the boxed pages. As complete files, they
are also available in the \GLOBES\ software package.

@@

\paragraph{Note:} All examples for application software in C do require
a C++ compiler to be properly compiled. For pedagogical reasons, variable
declarations are done at that place where the variable is needed for the
first time, which is at variance with C syntax but not with C++ syntax.
That is the only way in which the examples deviate form ISO C. Moreover
the actual numerical values of the results of the examples may be different
from the ones in this manual.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART III: Module Support
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% not yet!
%\input{part3}

@@

We would like to thank Martin Freund, who wrote the very first
version of a three-flavour matter profile treatment many years ago.
PH is especially thankful for the invaluable advice of Thomas Fischbacher on
many design issues.
In addition, we would like to thank
Mark Rolinec for his help to translate the experiment descriptions
into \AEDL, and for creating the illustrations in this manuscript.
Finally, thanks to all the people who have been pushing this project
for many years, to the ones who have been continuing asking for the
publication of the software, and the referees of several of our
papers for suggestions which lead to improvements in the software.

This work and the development of \GLOBES\ have been supported by:
\begin{itemize}
\item Physik Department der Technischen Universit\"at M\"unchen
\item Max-Planck-Institut f\"ur Physik
\item Sonderforschungsbereich 375 f\"ur Astro-Teilchenphysik\\
der Deutschen Forschungsgemeinschaft
\item Studienstiftung des Deutschen Volkes
\item Institute for Advanced Study, Princeton
\item Physics Department of the University of Wisconsin, Madison
\end{itemize}

>>>>

\newcommand{\glbxxx}{hep-ph/0407333}
\newcommand{\glbxxxtwo}{hep-ph/XXYYZZZ}
\newcommand{\capdef}{}
\newcommand{\mycaption}[2][\capdef]{\renewcommand{\capdef}{#2}%
        \caption[#1]{{ %\itshape
 {\footnotesize #2}}}}
\makeatletter
\renewcommand{\fnum@table}{\textbf{\tablename~\thetable}}
\renewcommand{\fnum@figure}{\textbf{\figurename~\thefigure}}
\makeatother

@@

\begin{center}
{\large Patrick Huber$^a$, Joachim Kopp$^b$, Manfred Lindner$^b$, \\ Mark Rolinec$^c$, Walter Winter$^d$}
\end{center}

@@

\begin{center}
Version from \today\ for \GLOBES\ 3.0
\end{center}

@@

\vspace*{0.1cm}

{\em\small $^{\mathrm{a}}$%
       University of Wisconsin,  Physics Department,
       1150 University Av., Madison, WI 53706, USA}

{\em\small $^{\mathrm{b}}$%
       Max--Planck--Institut f\"ur Kernphysik, 
       Postfach 10~39~80, D--69029 Heidelberg, Germany}

{\em\small $^{\mathrm{c}}$%
       Technische Universit\"at M\"unchen,
       Institut f\"ur Theoretische Physik, Physik--Department,\\
       James--Franck--Strasse, D--85748 Garching, Germany}

{\em\small $^{\mathrm{d}}$%
       Universit\"at W\"urzburg, 
       Lehrstuhl f\"ur theoretische Physik II, \\
       Institut f\"ur theoretische Physik und Astrophysik, 
       Am Hubland,
       D-97074 W\"urzburg, Germany}

@@

\clearpage
\thispagestyle{empty}
\bigskip
\begin{quote}
    Copyright \copyright  2004--2007  The GLoBES Team.
    Permission is granted to copy, distribute and/or modify this document
    under the terms of the GNU Free Documentation License, Version 1.2
    or any later version published by the Free Software Foundation;
    with the invariant Sections ``Terms of usage of \GLOBES" 
    and ``Acknowledgments'', no Front-Cover Texts, and no Back-Cover Texts.
    A copy of the license is included in the section entitled "GNU
    Free Documentation License".
\end{quote}
\bigskip

@@

On the experiment definition side, either built-in neutrino fluxes
(\eg, neutrino factory, $\beta$-Beam) or arbitrary fluxes can be used. Similarly,
arbitrary cross sections, energy dependent efficiencies, the
energy resolution function, the considered oscillation channels, 
backgrounds, and many other features can be specified. 
For the systematics, energy
normalization and calibration errors can be simulated in a straightforward way,
or the systematics can be completely user-defined (Version 3.0 and higher). Note that
the energy ranges and windows, as well as the bin widths can be
(almost) arbitrarily chosen, which means that variable bin
widths are allowed. Together with \GLOBES\ comes a number of
pre-defined experiments in order to demonstrate the capabilities
of \GLOBES\ and to provide prototypes for new experiments.

@@

\GLOBES\ is developed for academic use. Thus, the \GLOBES\ Team would
appreciate being given academic credit for it. Whenever you use \GLOBES\
to produce a publication or a talk indicate that you have used \GLOBES\ and
please cite the following references~\cite{globes_paper,globes_paper_two}
\begin{quote}
P. Huber, M. Lindner and W. Winter\\
Simulation of long baseline neutrino oscillation experiments with \GLOBES\\
Comput. Phys. Commun. 167 (2005) 195, arXiv:\glbxxx.   
\end{quote}

\begin{quote}
P. Huber, J. Kopp, M. Lindner, M. Rolinec, and W. Winter\\
TO BE DEFINED, NEEDS TO BE UPDATED [WW]\\
arXiv:\glbxxxtwo.   
\end{quote}
but \emph{not} this manual. This manual itself is not a scientific 
publication and will not be submitted to a scientific journal. 
It will evolve during time since it is intended for 
regular revision. Besides that, many of the data which are used by \GLOBES\ 
and distributed together with it should be properly referenced. 
For details see below.

@@

\index{norm}{Referencing!data in \GLOBES}
\GLOBES\ wouldn't be useful without having high quality input data.
Much of these input data have been published elsewhere and the authors
of those publications would appreciate to be cited whenever their work
is used. It is solely the user's responsibility 
to make sure that he understands where the input material for \GLOBES\ comes
from and if additional work has to be cited in addition to the 
\GLOBES\ papers~\cite{globes_paper,globes_paper_two}. To assist with this task, we provide  the necessary information for the data coming along together with \GLOBES.

@@

\chapter*{What is new in \GLOBES ?}

TO BE UPDATED [WW].

Here we briefly summarize the main changes of the new GLoBES version. For details, please
refer to the respective parts of the manual. Please note that any new \GLOBES\ version is
compatible with older versions, \ie, old application software and AEDL files should also run with the new version. EXCEPTION: VERSION NUMBER IN AEDL REQUIRES? [WW] However, some function names and features will
evolve during time, which means that outdated features may not be documented anymore.

\section*{Version 3.0}
\index{norm}{Version 3.0}

Here comes a summary of the most important changes in this version for older users of GLoBES.

\subsection*{New features}

\begin{itemize}
 \item
  User-defined systematics to simulate experiments with correlated systematics \etc\ (such as
reactor experiments); see \Sec~?
 \item
  User-defined priors to include arbitrary external information in the $\chi^2$ before
marginalization over the oscillation parameters (such as to prevent minimizer from
running into the wrong octant \etc); see \Sec~?
\item
  Non-standard physics support; see \Sec~?
\item
  Beta beam fluxes available as built-in fluxes; see \Sec~?
\item
  Enhanced support for parallelization, such as Condor; see \Sec~?
\item
  Updated AEDL files, new beta beam files; see \Sec~?
\item
 Faster probability engine
\item
  ...
\end{itemize}

\subsection*{Major changes}

Most of the modifications below should not require that old pieces of software be changed. 
However, the following changes are necessary:
\begin{itemize}
\item
 AEDL requires now that {\tt \$version} be used to define the minimum version number this AEDL
file is to be used with. With this requirement one can easily avoid that new AEDL
files with new features be used with old versions of \GLOBES\ which may not recognize these features.
\item
 Some of the earlier AEDL files have been updated, changed names, or have been removed. In addition, new files have been added. Although old AEDL files will run as usual for compatibility, they will not be
supported by the \GLOBES\ team anymore. You should make sure to keep these files when updating \GLOBES . 
\item
 ...
\end{itemize}
The following changes should not affect older software:
\begin{itemize}
\item
 {\tt glbChi...} functions renamend for consistency
\item
 {\tt glbChiSetStartingValues} and {\tt glbChiGetStartingValues} renamed into 
{\tt glbChiSetCentralValues} and {\tt glbChiGetCentralValues}
\item
 Systematics concept changed, concept of error dimensions removed
\item
 Some functions changed (such as for direct rate access)
\item
 ..
\end{itemize}

@@

The application 
software is, except from some example files, not part of \GLOBES , since
the evaluation of the experiment performance is often a matter of taste
and definition. In addition, the algorithms depend, especially for
high-precision instruments, very much on the oscillation parameters.
In general, it is quite simple to simulate superbeams and reactor
experiments. However, because of the more complicated topology, the
simulation of neutrino factories is much more difficult. In order
to demonstrate some of these difficulties, we present in this manual mainly
examples with neutrino factories. These examples can be found in
\Part~\ref{part:1} within the boxed pages. As complete files, they
are also available in the \GLOBES\ software package.

@@

\paragraph{Note:} All examples for application software in C do require
a C++ compiler to be properly compiled. For pedagogical reasons, variable
declarations are done at that place where the variable is needed for the
first time, which is at variance with C syntax but not with C++ syntax.
That is the only way in which the examples deviate form ISO C. Moreover
the actual numerical values of the results of the examples may be different
from the ones in this manual.  (IS THAT STILL SO? [WW])

@@

@@

We would like to thank Martin Freund, who wrote the very first
version of a three-flavor matter profile treatment many years ago,
and Thomas Schwetz, who has been pushing the software to the edge in the 
past few years. Furthermore, we would like to thank Toshihiko Ota and WHO ELSE? [WW]
for using and testing unpublished new features of the software.
PH is especially thankful for the invaluable advice of Thomas Fischbacher on
many design issues in the early stage of the project.
% In addition, we would like to thank
% Mark Rolinec for his help to translate the experiment descriptions
% into \AEDL, and for creating the illustrations in this manuscript.
Finally, thanks to all the people who have been pushing this project
for many years, to the ones who have been continuing asking for the
publication of the software, and the referees of several of our
papers for suggestions which lead to improvements in the software.

This work and the development of \GLOBES\ have over the years been supported by:
\begin{itemize}
\item Technische Universit\"at M\"unchen [All authors]
\item Max-Planck-Institut f\"ur Physik, M\"unchen [PH]
\item Sonderforschungsbereich 375 f\"ur Astro-Teilchenphysik\\
der Deutschen Forschungsgemeinschaft [All authors]
\item Studienstiftung des Deutschen Volkes [JK, WW]
\item Institute for Advanced Study, Princeton [WW]
\item W. M. Keck Foundation [WW]
\item National Science Foundation [WW]
\item University of Wisconsin, Madison [PH]
\item Max-Planck-Institut f\"ur Kernphysik, Heidelberg [JK, ML]
\item Emmy Noether-Programm der Deutschen Forschungsgemeinschaft [WW]
\item Universit\"at W\"urzburg [WW]
\end{itemize}
